On the marginal value of swimming in woodland caribou

### 2021-05-25 paper/manuscript.Rmd
P1 and P2

Movement is costly. Even for adept swimmers like caribou (*Rangifer tarandus*), which have hollow hair
shafts that aid in flotation, the energetic expenditure of swimming for terrestrial mammals is
markedly higher than walking or running (Miller and Gunn 1985). The drivers that promote the decision
to swim for terrestrial mammals, like caribou, are often related to the profitability of forage
(Miller et al. 1977) and the risks of predation (Jeffery et al. 2007, Jordan et al. 2010). The Fogo
Island archipelago, off the coast of Newfoundland, Canada, is home to approximately 300 woodland
caribou, descendants of 26 individuals introduced from the Island of Newfoundland to Fogo Island
between 1964–67 (Bergerud and Mercer 1989). During fieldwork on May 30, 2017, we observed an unmarked
adult male caribou swim between two smaller islands (Figure 1) a distance of 470 m which took
approximately 9 minutes. Given that swimming is energetically expensive, we ask how often do caribou
swim between islands, how long do they occupy islands, and ultimately what motivates swimming? Here,
we frame swimming and island residency as a model of patch use as described by the marginal value
theorem (Charnov 1976) within the context of the forage limitation hypothesis. For caribou, the forage
limitation hypothesis is the constraints on food availability associated with changes in abundance
through time (Klein 1991, Schaefer et al. 2016). Here, we integrate our understanding of forage
limitation in caribou populations by using island residency as a proxy for time in patch and island
size as a proxy for energy intake in a patch (Figure 2).

Forage limitation and over-grazing is a concern for caribou populations and can lead to reduced female
reproductive success (Klein 1991, Schaefer et al. 2016). Forage scarcity has been proposed as a
potential reason for caribou moving between arctic islands on the sea-ice in winter (Miller et al.
1977). We submit that testing the marginal value theorem could improve our understanding of forage
limitation. The marginal value theorem posits that animals will forage optimally by depleting the
resources in a given patch (i.e. island) before moving to a new patch (Charnov 1976). Depletion of
resources on small islands could therefore mirror patterns of over-grazing and forage limitation that
occur for caribou herds at higher densities in Newfoundland (Schaefer et al. 2016) and elsewhere in
their range (Klein 1991).

>>>>

Movement is costly. Even for adept swimmers like caribou (*Rangifer tarandus*), which 
have hollow hair shafts that aid in flotation, the energetic expenditure of swimming 
for terrestrial mammals is markedly higher than walking or running (Miller and Gunn 1985). 
The drivers that promote the decision to swim for terrestrial mammals, like caribou, 
are often related to the profitability of forage (Miller et al. 1977) and the risks 
of predation (Jeffery et al. 2007, Jordan et al. 2010). The Fogo Island archipelago, 
off the coast of Newfoundland, Canada, is home to approximately 300 woodland caribou, 
descendants of 26 individuals introduced from the Island of Newfoundland to Fogo Island 
between 1964–67 (Bergerud and Mercer 1989). During fieldwork on May 30, 2017, we 
observed an unmarked adult male caribou swim between two smaller islands (Figure 1), 
a distance of 470 m which took approximately 9 minutes. Given that swimming is 
energetically expensive, we ask how often do caribou swim between islands, how 
long do they occupy islands, and ultimately what motivates swimming? Here, we 
frame swimming and island residency as a model of patch use as described by the 
marginal value theorem (Charnov 1976) within the context of the forage limitation 
hypothesis. For caribou, the forage limitation hypothesis refers to the 
constraints on food availability associated with changes in abundance (Klein 1991, 
Schaefer et al. 2016). Here, we integrate our understanding of forage limitation in
caribou populations by using island residency as a proxy for time in a patch and 
island size as a proxy for energy intake in a patch (Figure 2).

Forage limitation and over-grazing can reduce reproductive success of female caribou (Klein 1991, Schaefer et al. 2016). Forage scarcity has been proposed as a potential reason for caribou moving between arctic islands on the sea-ice in winter (Miller et al. 1977). We submit that testing the marginal value theorem could improve our understanding of forage limitation. The marginal value theorem posits that animals will forage optimally by depleting the resources in a given patch (i.e. island) before moving to a new patch (Charnov 1976). Depletion of resources on small islands could therefore mirror patterns of over-grazing and forage limitation that occur for caribou herds at higher densities in Newfoundland (Schaefer et al. 2016) and elsewhere in their range (Klein 1991).
